---
title: "Week 07 - Time-Series Cross-Sectional Analysis and Causal Inference with Observational Data"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: false
    css: "css/learnr-theme.css"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(learnrhash)
library(texreg)
library(broom)
library(highlight)
library(gradethis)
library(texreg)
library(tidyverse)
library(qmir22)

gradethis:: gradethis_setup()

data(df_rai)
data(eu_accession)
```

## Time-series cross-sectional analysis

Welcome to this week's learnr-exercises (the last ones for this semesters)! 

In the first part of this tutorial, you will run a pooled model, a fixed-effects model, compare the results, and predict values for Japan and Uruguay! Before we start modeling, though, let's repeat some key contents of the lecture. 

```{r q1, echo = FALSE}
question("Why does it make sense to use a TSCS model?",
  answer("To adress concerns about omitted variable bias", correct = T),
  answer("To account for potential exogenous shocks", correct = T),
  answer("To deal with NAs in our data set"),
  answer("To incorporate unit-specific, time-varying characteristics in our model"),
  random_answer_order = F,
  allow_retry = T,
  incorrect = "Hint: You may want to check the lecture materials."
)
```

```{r q2, echo = FALSE}
question("What is commonly *not* constant within units over time?",
  answer("The electoral system", message = "How often has the electoral system in your home-country changed in your lifetime?"),
  answer("The unemployment rate", correct = T),
  answer("Culture", message = "Though you may argue that culture is everchanging and subject to the understanding of the individual, for an FE-regression we may assume that it is somewhat constant over time."),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: Think about how much variation there is in each answer option."
)
```

```{r q3, echo = FALSE}
question("When may we want to use a two-way FE model?",
  answer("When we want to test two competing explanations", message = "We can do this with or without a fixed effect!"),
  answer("When we want to account for two variables that are constant over time", message = "Two variables that are constant over time, but differ from unit to unit would be covered with a unit-fixed effect."),
  answer("When we want to account for differences between units that are constant over time and for exogenous shocks that affect all units simultaneously", correct = T),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: You may want to check the lecture materials."
)
```

```{r q4, echo = FALSE}
question("What happens when we use a unit-fixed effect?",
  answer("It changes the intercept for each unit", correct = T),
  answer("It changes the slope for each unit"),
  answer("It runs a separate model for each unit"),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: Check the lecture materials."
)
```


## Regression exercises (TSCS)

Góméz-Díaz and Góméz-Díaz (2020) have argued that prolonged membership in the European Union leads to an increase in regional authority in different member-states. However, they also include a variable for the duration of World Bank membership. In the following, we will focus on the effect of World Bank membership on sub-national authority. In a final step we will calculate predicted values for Japan and Uruguay.

In the following exercises you will need these variables:

| Variable name |  Description
|:--------------|:-------------
| `n_RAI` | The dependent variable. Measures the regional authority
| `length_EU` |  Duration of EU membership in years
| `length_world` | Duration of World Bank membership in years
| `democ` | Polity2 democracy index
| `gdp_capita` | Gross domestic product per capita
| `population_size` | Size of population
| `nationalization` | Nationalization of the party system
| `country_name` | Name of the country


### Preparing the data and running our first model.


Let's take a look at our data `df_rai`. In the first step, we want you to run a pooled model using the following variables: `n_RAI`, `length_EU`, `length_world`, `democ`, `gdp_capita`, `population_size`, and `nationalization`.


```{r exercise1, exercise = TRUE}
# Let's run our first model and call it pooled_model

```

```{r q5, echo = FALSE}
question_text("What is the coefficient for the duration of World Bank membership (`length_world`) in the pooled model? Round the coefficient to two digits after the decimal.",
  answer("-0.04", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: Have you rounded correctly?"
)
```


Now run the same model, but include a fixed-effect for the unit. Our unit is the country (`country_name`)`. Let's call this model "fe_model".
```{r exercise2, exercise = TRUE}

```


```{r q6, echo = FALSE}
question_text(" What is the unit-specific intercept for Albania? Round the coefficient to two digits after the decimal.",
  answer("0.86", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: is the fixed-effect correctly specified?"
)
```

```{r q7, echo = FALSE}
question_text("What is the coefficient for the duration of World Bank membership  (`length_world`) in the fixed-effects model? Round the coefficient to two digits after the decimal.",
  answer("0.03", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: is the fixed-effect correctly specified?"
)
```



```{r q8, echo = FALSE}
question_text("If you wanted to present your results in a table using `screenreg()`, but wanted to exclude the country FE's from the table, what argument would you use?",
  answer("omit.coef", correct = T),
  answer("omit.coef = country_name", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: Take a look at the arguments of the screenreg()-function."
)
```


```{r q9, echo = FALSE}
question("Comparing the coefficients from the pooled model and from the FE model, what do you conclude?",
         answer("Due to omitted variable bias our estimate we would have expected a negative relationship between the duration of World Bank membership. After accounting for the unobserved, constant variation between countries, we were able to show that the relationship is, in fact, positive and statistically significant.", correct = T),
         answer("Accounting for the unobserved, constant variation between countries has shown that there is no effect of World Bank membership on regional authority"),
         answer("We have different numbers of observation and therefore the coefficients differ.", message = "Take a look at the number of observations included in each model. Are they the same or do they differ?"),
         answer("The coefficients are the same.", message = "If the coefficients are the same, you may want to retrace your steps and check your code whether there is an error."),
         random_answer_order = TRUE,
         allow_retry = T
)
```

### Prediction

Remember from the last sessions that predicted values can help us to communicate the results of our analysis. In the following section, the model has already been specified. Please calculate predicted values and confidence intervals for the maximum duration of membership in the **World Bank** for Uruguay and Japan. Compare the predictions. In which of the two countries is the predicted value higher? Don't forget about the confidence intervals.

```{r exercise3, exercise = TRUE}
fe_model <-  lm(n_RAI ~ length_EU + length_world + democ + gdp_capita + population_size + nationalization + country_name-1, data = df_rai)

```

```{r q10, echo = FALSE}
question("Comparing the predicted values, in which country is the prediction for sub-national authority higher?",
  answer("Japan", message = "Keep in mind that confidence intervals are important!"),
  answer("Uruguay", message = "Keep in mind that confidence intervals are important!"),
  answer("Unclear, because the confidence intervals overlap.", correct = T),
  random_answer_order = TRUE,
  allow_retry = T)
```

In a final step, we want you to calculate predicted values for minimum and maximum values of World Bank membership for Japan and compare them.

```{r exercise4-setup}
fe_model <-  lm(n_RAI ~ length_EU + length_world + democ + gdp_capita + population_size + nationalization + country_name-1, data = df_rai)
```

```{r exercise4, exercise = TRUE,  exercise.setup = "exercise4-setup"}


```

```{r q11, echo = FALSE}
question_text("What is the difference in predicted values for minimum and maximum World Bank membership for Japan? Round to two digits after the decimal",
  answer("2.10", correct = T),
  answer("2.1", correct = T),
  allow_retry = T,
  incorrect = "Did you correctly specify the scenarios?")
```

## Causal inference with observational data

The second part of this tutorial revolves around **how accession to the European Union (EU) affects citizens' anti-authoritarian attitudes**. Before you begin to work with the data, you will have to answer some questions to revise what you have learned this week.


### Questions

```{r q1-causal, echo = FALSE}
question("Why is it so difficult to make causal statements with observational data?",
  answer("The observed difference in outcomes might be caused by changes in unobserved confounders instead of the treatment.", correct = T),
  answer("Treatments do not have an effect because of unobserved confounders.", 
         message = "Think about the role of confounders in analyses of observational data. The treatment may work, but we do not know what influence the confounders have had which biases our estimate of the treatment effect. Have a look at the lecture material again."),
  answer("Observational data does not tell us which unit is from the treatment and the control group.", 
         message = "Whether a unit belongs to the treatment or the control group depends on the value it takes for the main independent variable, the treatment. Have a look at the lecture material again."),
  random_answer_order = TRUE,
  allow_retry = TRUE
)
```

###

```{r q2-causal, echo = FALSE}
question("What is the main identifying assumption of difference-in-differences designs?",
  answer("The change in the outcome variable in the treatment group in absence of treatment is equal to the change in the control group over time.", 
         correct = T),
  answer("The outcome for the treatment group in absence of the treatment is identical to the outcome of the control group.", 
         message = "Not quite. While we use the outcome of the control group when calculating the difference-in-differences estimate, we do not assume that the treatment group's outcome is identical to the control group's outcome -- only the relative change is assumed to be similar. Have a look at the lecture material again."),
  answer("The average causal effect is the difference in the treatment group's outcome before and after the treatment.", 
         message = "In a before-and-after design, we calculate the different in the treatment group's outcome before and after the treatment. However, this does not necessarily allow us to interpret the difference as causal as there might be time-varying confounders. Have a look at the lecture material again."),
  random_answer_order = TRUE,
  allow_retry = T
)
```

### Data

Take a look at the data set for this learnr tutorial (`eu_accession`). As always, the data set is already loaded into the environment and contains the following variables:

| Variable name |  Description
|:--------------|:--------------------------------
| `wave` | Time frame indicating the period of observation
| `countryname` |Country name
| `cntrymean_antiauth` | Anti-authoritarianism score
| `EU` | EU membership (1 = EU member, 0 = no EU member)

Use the code chunk below, to explore the data set.

```{r inspect-data-causal, exercise=TRUE}

```

<div id="inspect-data-hint">
**Hint:** You may want to use functions like `glimpse()`, `head()`, `names()`, `dim()`, `nrow()` or `ncol()` or some of the **dplyr** functions we learned about in the course.
</div>

### Exercises

As you will explore changes in anti-authoritarian attitudes in the `eu_accession` data, start by calculating the average degree of anti-authoritarianism (`cntrymean_antiauth`) among EU members and non-EU members (`EU`) in the 2008-2010 time period (`wave`). Please round to the second digit after the decimal.

```{r exercise1-causal-sol, eval=FALSE, include=FALSE}
eu_accession %>% filter(wave == "2008-2010" & EU == 1) %>% summarise(mean = round(mean(cntrymean_antiauth), 2))
eu_accession %>% filter(wave == "2008-2010" & EU == 0) %>% summarise(mean = round(mean(cntrymean_antiauth), 2))

# or
eu_accession %>% filter(wave == "2008-2010") %>% group_by(EU) %>% summarise(mean = round(mean(cntrymean_antiauth), 2))
```


```{r exercise1-causal, exercise = TRUE}

```

```{r q3-causal, echo = FALSE}
question_text("What is the average degree of anti-authoritarianism among EU members in 2008-2010?",
  answer("75.69", correct = T),
  answer("75.7", correct = T),
  answer("75.70", correct = T),
  allow_retry = TRUE,
  incorrect = "Have you selected the relevant observations with \"filter()\" and calculated the average across all those observations using the \"mean()\" function?"
)
```

```{r q4-causal, echo = FALSE}
question_text("And what about the average degree of anti-authoritarianism among non-EU members in the same time frame?",
  answer("67.2", correct = T),
  answer("67.18", correct = T),
  answer("67.20", correct = T),
  allow_retry = TRUE,
  incorrect = "Have you selected the relevant observations with \"filter()\" and calculated the average across all those observations using the \"mean()\" function?"

)
```

###

In the next few tasks, you will calculate the causal effect of EU membership on anti-authoritarian attitudes of citizens using a difference-in-differences design. In particular, you will calculate the causal effect of EU accession on anti-authoritarian values for Romania (= *treatment unit*) compared to Albania (= *control unit*). Based on the parallel trends assumption, you assume that anti-authoritarian attitudes would have developed similarly in both countries in the absence of Romania's EU accession to interpret the estimate as causal.

Calculate the change in anti-authoritarian values for Romania between 2008-2010 and 1999-2004. Round to two digits after the decimal. 

```{r exercise2-causal-sol, eval=FALSE, include=FALSE}
eu_accession %>% 
  # Filter for treatment unit
  filter(countryname == "Romania") %>% 
  # Filter for pre- and post-treatment period
  filter(wave == "1999-2004" | wave == "2008-2010") %>% 
  # Pivot data into wide format to facilitate calculations
  pivot_wider(names_from = "wave",
              values_from = "cntrymean_antiauth") %>%
  # Calculate difference within unit
  mutate(difference_within = round(`2008-2010` - `1999-2004`, 2))
```

```{r exercise2-causal, exercise = TRUE}

```

```{r q5-causal, echo = FALSE}
question_text("Enter the change in anti-authoritarian values for Romania between 2008-2010 and 1999-2004, rounded to two digits after the decimal, here.",
  answer("0.32", correct = T),
  allow_retry = TRUE,
  incorrect = "Following what we have done in the lab, calculate the values of \"cntrymean_antiauth\" for Romania for these two waves and subtract them from each other."
)
```

###

Now, calculate the change in anti-authoritarian values for Albania between 2008-2010 and 1999-2004. Round to two digits after the decimal.

```{r exercise3-causal-sol, eval=FALSE, include=FALSE}
eu_accession %>% 
  # Filter for control unit
  filter(countryname == "Albania") %>% 
  # Filter for pre- and post-treatment period
  filter(wave == "1999-2004" | wave == "2008-2010") %>% 
  # Pivot data into wide format to facilitate calculations
  pivot_wider(names_from = "wave",
              values_from = "cntrymean_antiauth") %>%
  # Calculate difference within unit
  mutate(difference_within = round(`2008-2010` - `1999-2004`, 2))
```

```{r exercise3-causal, exercise = TRUE}

```

```{r q6-causal, echo = FALSE}
question_text("Enter the change in anti-authoritarian values for Albania between 2008-2010 and 1999-2004, rounded to two digits after the decimal, here.",
  answer("-4.77", correct = T),
  allow_retry = TRUE,
  incorrect = "Following what we have done in the lab, calculate the values of \"cntrymean_antiauth\" for Albania for these two waves and subtract them from each other."
)
```

###

Next, calculate the average treatment effect for the treated, i.e. the average treatment effect of EU accession for Romania. Based on the two previous exercises you already have all the information you need: the difference in anti-authoritarian values over time for Romania and the difference in anti-authoritarian values over time for Albania. Alternatively, the within-unit differences are already stored in a variable called `difference_within` in two objects called `romania` and `albania`. You can access the values with `romania$difference_within` and `albania$difference_within`.

```{r exercise4-causal-setup}
romania <- eu_accession %>% 
  # Filter for treatment unit
  filter(countryname == "Romania") %>% 
  # Filter for pre- and post-treatment period
  filter(wave == "1999-2004" | wave == "2008-2010") %>% 
  # Pivot data into wide format to facilitate calculations
  pivot_wider(names_from = "wave",
              values_from = "cntrymean_antiauth") %>%
  # Calculate difference within unit
  mutate(difference_within = round(`2008-2010` - `1999-2004`, 2))

albania <- eu_accession %>% 
  # Filter for control unit
  filter(countryname == "Albania") %>% 
  # Filter for pre- and post-treatment period
  filter(wave == "1999-2004" | wave == "2008-2010") %>% 
  # Pivot data into wide format to facilitate calculations
  pivot_wider(names_from = "wave",
              values_from = "cntrymean_antiauth") %>%
  # Calculate difference within unit
  mutate(difference_within = round(`2008-2010` - `1999-2004`, 2))
```

```{r exercise4-causal-sol, eval=FALSE, include=FALSE}
romania$difference_within - albania$difference_within
```


```{r exercise4-causal, exercise = TRUE, exercise.setup = "exercise4-setup"}


```

```{r q8-causal, echo = FALSE}
question_text(
  "What is the average treatment effect of EU accession for Romania? Round to two digits after the decimal",
  answer("5.09", correct = T),
  allow_retry = TRUE,
  incorrect = "The average treatment effect for the treated (in this case, Romania) is the difference between the change in outcomes for the treatment unit (Romania) and the change in outcomes for the control unit (Albania)."
)
```

```{r q9-causal, echo = FALSE}
question("Does this mean that anti-authoritarian attitudes in Romania have increased or decreased due to EU accession?",
  answer("increased", correct = T),
  answer("decreased"),
  random_answer_order = FALSE,
  allow_retry = T,
  incorrect = "Think about what the treatment effect means. Is it positive or negative?"
)
```

###

In a final step, use a difference-in-differences regression model to analyze the effect of EU accession on anti-authoritarianism. The subset of the data that you will use as well as the treatment time and treatment group are already specified in the code chunk below. All you need to do is to add the regression model.

```{r exercise5-causal-sol, eval=FALSE, include=FALSE}
lm(cntrymean_antiauth ~ treatment_time*treatment_group, data = did_data) %>%
  broom::tidy() %>%
  filter(term == "treatment_time:treatment_group")
```

```{r exercise5-causal, exercise=TRUE}
did_data <- eu_accession %>% 
  tibble() %>%
  filter(countryname %in% c("Romania", "Albania", "Finland", "Ukraine", "Switzerland")) %>%
  mutate(treatment_time = ifelse(wave == "2008-2010", 1, 0),
         treatment_group = ifelse(countryname == "Romania", 1, 0))
```

```{r q10-causal, echo=FALSE}
question("Which countries form the control group?",
  answer("Albania, Finland, Ukraine and Switzerland", correct = T),
  answer("Romania", 
         message = "Romania is the treatment unit as it has become an EU member between the two time points."),
  answer("Albania", 
         message = "Albania is one of the control units, but not the only one. Have a look at the other countries that were selected."),
  random_answer_order = TRUE,
  allow_retry = TRUE,
  incorrect = "Inspect the code chunk above."
)
```

```{r q11-causal, echo=FALSE}
question_text("The average treatment effect of EU accession  based on the regression model is... (rounded to the second digit after the decimal)",
  answer("-8.23", correct = T),
  allow_retry = TRUE,
  incorrect = "Have you specified the data and the model correctly? Have you rounded to two digits after the decimal?"
)
```



## Submit

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(default_ui(url = ""))
```
