---
title: "Week 06 - Interaction Models and Logistic Regression Analysis"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: false
    css: "css/learnr-theme.css"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(learnrhash)
gradethis::gradethis_setup()
library(highlight)
library(tidyverse)
library(broom)
library(texreg)
library(margins)
library(qmir22)

data(df_brexit) 
data(df_opposition)

```

## Modeling interaction effects

You will continue working with the data on opposition party behavior in European national parliaments by Tuttnauer (2018). This week, you are asked to reproduce some of the results from the paper with a focus on modeling and interpreting interaction effects. 

Please start with reproducing model 6 from page 290 in Tuttnauer (2018). The dependent variable is, as in previous weeks, `conflict_p`. The model includes an interaction of `seatshare` and government experience, (`ge`) as the key explanatory terms.  Moreover, the model controls for the frequency of recorded votes (`frequency`), whether or not an opposition is on the ideological side of the government (`govside`), and the ideological range of the government coalition (`govrange`). 

Don't forget to include the other control variables from last week:

* `abs_lrdiff`: Ideological distance from the government as the absolute difference between a party's left-right position and the size-weighted average position of the government coalition.
* `inst_power`:  Institutional power of the opposition measuring the strength of the committee system on a scale from 0 to 7.
* `y_openness`: Average number of distinct government compositions in the past 10 years.
* `importance`: Index ranging between 0 (low-importance parliament) and 1 (high-importance parliament).
* `govtype_mod`: A dummy (=1) for surplus coalitions, in contrast with minimal-winning ones (=0).
* `record_threshold`:The percentage of MPs needed to call for a recorded vote (ranging between 0 and 0.3).

**Please specify and run model 6 from Tuttnauer (2018) and report the constitutive term of party size as well as the interaction term of party size and government experience. Round to two digits after the decimal.**

```{r lm_1_sol, eval=FALSE, echo = FALSE}
model <- lm(conflict_p ~ seatshare*ge + abs_lrdiff + govside + inst_power + 
                  y_openness + importance + govrange + govtype_mod + record_threshold + frequency,
                data = df_opposition)

model %>% tidy() %>% filter(term == "seatshare")
```

```{r lm_1, exercise = TRUE}

```

```{r lm_1_q, echo = FALSE}
question_text("Enter the value of the constitutive term of  seat share:",
  answer("1.92", correct = T),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session. Make sure that you included all variables in the model."
)
```

```{r lm_2_q, echo = FALSE}
question_text("Enter the value of the interaction term of seat share and ge:",
  answer("-3.21", correct = T),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session. Make sure that you included all variables in the model."
)
```

## Interpreting interaction effects

Interpreting interaction effects can be tricky. As discussed in class, there are two main approaches: We can look at the marginal effect of a variable conditional on values of the moderating variable (as Tuttnauer did in Figure 2 in his paper) or we can calculate predicted values to generate meaningful quantities of interest. 

### Marginal effects

**What is the marginal effect of seat share when government experience is at its minimum in comparison to when government experience is at its maximum (as well as the corresponding 95% confidence interval for the marginal effect estimates)?**

*Hint: You need the `margins()` function from the `margins` package for this exercise. In order to inspect the output from the `margins()` function, the `summary()` function may be helpful.*

*You can re-run the regression model in the chunk below. Alternatively, the correct regression output is already stored in an object called `out_ia`.*

```{r margins_sol, echo = FALSE, eval = FALSE}
out_ia <- lm(conflict_p ~ seatshare*ge + abs_lrdiff + govside + inst_power + 
                  y_openness + importance + govrange + govtype_mod + record_threshold + frequency,
                data = df_opposition)

mar <- out_ia %>%
  margins(
    # marginal effect of which variable:
    variables = "seatshare",
    # conditional on which values:
    at = list(ge = c(min(df_opposition$ge), max(df_opposition$ge)))
  ) 

summary(mar) 
```

```{r margins-setup}
out_ia <- lm(conflict_p ~ seatshare*ge + abs_lrdiff + govside + inst_power + 
                  y_openness + importance + govrange + govtype_mod + record_threshold + frequency,
                data = df_opposition)
```

```{r margins, exercise = TRUE, exercise.setup = "margins-setup"}

```


```{r margeff_min, echo = FALSE}
question_text("Enter the value of the (average) marginal effect of party size when government experience is at its minimum (rounded to two digits after the decimal):",
  answer("1.92", correct = T),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session, in particular, the margins() function."
)
```

```{r margeff_max, echo = FALSE}
question_text("Enter the value of the (average) marginal effect of party size when government experience is at its maximum (rounded to two digits after the decimal):",
  answer("-1.29", correct = T),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session, in particular, the margins() function."
)
```

```{r margeff_low, echo = FALSE}
question_text("Enter the value of the lower bound of the 95% confidence interval of the marginal effect of party size when  government experience is at its minimum (rounded to two digits after the decimal):",
  answer("0.95", correct = T),
  answer("-2.71", correct = F, message = "Make sure you report the result for the marginal effect when government experience is at its minimum."),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session, in particular, the margins() function."
)
```

```{r margeff_high, echo = FALSE}
question_text("Enter the value of the upper bound of the 95% confidence interval of the marginal effect of party size when  government experience is at its maximum (rounded to two digits after the decimal):",
  answer("0.13", correct = T),
  answer("2.90", correct = F, message = "Make sure you report the result for the marginal effect when government experience is at its maximum."),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session, in particular, the margins() function."
)
```

**So what do we learn from the marginal effect estimation?**

```{r margeff_interpret, echo = FALSE}
question("The marginal effect of seat share...",
  answer("is positive and statistically significant at the 95% confidence level when government experience is at its minimum.", correct = T),
  answer("is negative and statistically significant at the 95% confidence level when government experience is at its maximum.", correct = F, message = "Have a look at the confidence interval for seat share when government experience is at its maximum. Does it include zero?"),
  answer("does not turn statistically significant at the 95% confidence level.", correct = F, message = "Statistical significance is conditional on the moderating variable."),
  answer("is statistically independent of government experience.", correct = F, message = "Statistical significance is conditional on the moderating variable."),
  random_answer_order = TRUE,
  allow_retry = T
)
```

### Predicted values

Estimating and reporting marginal effects is very common but can become complex, in particular, when we start dealing with non-linear regression models in a few weeks. Therefore, it is often recommendable to calculate meaningful predicted values for quantities of interest. Most of the times (and also in this case), researchers report predictions for the mean response (expected value) and not for future events. It allows them to ignore fundamental model uncertainty which is not of central concern when interpreting regression results.

**What is the predicted (expected) average opposition share (in percent) when party size is at its minimum in comparison to when party size is at its maximum if government experience is equal to one, holding everything else at the respective mean/median. Please also calculate the corresponding 95% confidence intervals. When submitting your results, please round them to two integer digits (e.g., 0.23, 4.5, 13, 60, 65)**

*Hint: The `predict()` function may be very useful for this exercise. Recall that govtype_mod and govside are binary variables when you specify a scenario of interest.*

*You can re-run the regression model in the chunk below. Alternatively, the correct regression output is already stored in an object called `out_ia`.*


```{r predict_sol, echo = FALSE, eval = FALSE}
model <- lm(conflict_p ~ abs_lrdiff + seatshare*ge + govside + inst_power + 
                  y_openness + importance + govrange + govtype_mod + record_threshold + frequency,
                data = df_opposition)

df_scenario <- tibble(
   seatshare = c(min(df_opposition$seatshare), max(df_opposition$seatshare))
  , ge = 1
  , abs_lrdiff = mean(df_opposition$abs_lrdiff)
  , govrange = mean(df_opposition$govrange)
  , govside = median(df_opposition$govside)
  , inst_power = mean(df_opposition$inst_power)
  , y_openness = mean(df_opposition$y_openness)
  , importance = mean(df_opposition$importance)
  , govtype_mod = median(df_opposition$govtype_mod)
  , record_threshold = mean(df_opposition$record_threshold)
  , frequency = mean(df_opposition$frequency)
)

round(predict(model, newdata = df_scenario, interval = "confidence")*100,0)

# manually
predict(model, newdata = df_scenario, se.fit = T) %>% as_tibble() %>%
  mutate(
    , fit = fit*100
    , se.fit = se.fit*100
    , lower = fit-1.96*se.fit
    , upper = fit+1.96*se.fit)

```

```{r predict-setup}
out_ia <- lm(conflict_p ~ seatshare*ge + abs_lrdiff + govside + inst_power + 
                  y_openness + importance + govrange + govtype_mod + record_threshold + frequency,
                data = df_opposition)
```


```{r predict, exercise = TRUE, exercise.setup = "lm_1"}

```


```{r predict_min, echo = FALSE}
question_text("Enter the value of the predicted average opposition share  (in percent) when government experience is equal to one and party size is at its minimum (rounded to two integer digits):",
  answer("67", correct = T),
  answer("67.42", correct = F, message = "Please round your results to two integer digits."),
  answer("67.4", correct = F, message = "Please round your results to two integer digits."),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session, in particular, the predict() function. Have you set all binary variables to their median (and not the mean)?"
)
```

```{r predict_max, echo = FALSE}
question_text("Enter the value of the predicted average opposition share (in percent)  when government experience is equal to one and party size is at its maximum (rounded to two integer digits):",
  answer("15", correct = T),
  answer("15.27", correct = F, message = "Please round your results to two integer digits."),
  answer("15.3", correct = F, message = "Please round your results to two integer digits."),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session, in particular, the predict() function."
)
```

```{r predict_low, echo = FALSE}
question_text("Enter the value of the lower bound of the 95% confidence interval (in percent) when government experience is equal to one and party size is at its minimum (rounded to two integer digits):",
  answer("41", correct = T),
  answer("42", correct = T),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session, in particular, the predict() function."
)
```

```{r predict_high, echo = FALSE}
question_text("Enter the value of the upper bound of the 95% confidence interval (in percent) when government experience is equal to one and party size is at its maximum (rounded to two integer digits):",
  answer("55", correct = T),
  answer("54", correct = T),
  allow_retry = T, incorrect = "Hint: Revisit the code from the lab-session, in particular, the margins() function."
)
```

**So what do we learn from the predictions?**

```{r predict_interpret, echo = FALSE}
question("The predicted average value for votes against the government...",
  answer("is unlikely to differ for opposition parties with maximum government experience conditional on their party size.", correct = T, message = "Due to the overlap of the confidence intervals, we have good reason to not exclude the possibility that there may be no difference in the expected value in the two specified scenarios. However, note that this is not an explicit statistical test whether there is a difference, but only a first (strong) indication."),
  answer("is likely to differ for opposition parties with maximum government experience conditional on their party size.", correct = F, message = "Have a closer look at the confidence intervals. What do you find?"),
  random_answer_order = TRUE,
  allow_retry = T
)
```

## Logistic regression analysis

For the second part of this week's tutorial, you are going to replicate some of the results of Sarah Hobolt's (2016) paper: \"The Brexit vote: a divided nation, a divided continent\". We will use different survey items to test hypotheses about the decision to vote "Leave" in the Brexit referendum. 

Before you start, take a look at the data set for this learnr tutorial (`df_brexit`). As always, the data set is already loaded into the environment and contains the following variables:


| Variable name |  Description
|:--------------|:--------------------------------
| `LeaveVote` | Whether a respondent intends to vote "Leave" (dummy)
| `edlevel` | Level of education (ordinal)
| `gender` | Gender (dummy)
| `age` | Age (continuous)
| `hhincome` | Household income (ordinal)
| `econPersonalRetro1` | Personal economic situation (ordinal)
| `EuropeanIdentity` | Level of identification with Europe (ordinal)
| `EnglishIdentity` | Level of identification with England (ordinal)
| `BritishIdentity` | Level of identification with Great Britain (ordinal)
| `female` | Female (dummy)

Use the code chunk below, to explore the data set. As you will note, a lot of the variables are **ordinal variables**. 


```{r inspect-data, exercise=TRUE}

```

<div id="inspect-data-hint">
**Hint:** You may want to use functions like `glimpse()`, `head()`, `names()`, `dim()`, `nrow()` or `ncol()` or some of the **dplyr** functions we learned about in the course.
</div>


### Knowledge Questions

```{r q1, echo = FALSE}
question("Why is it useful to rely on logistic regression models for our referendum vote variable?",
  answer("Because our dependent variable is binary and our predicted values may fall outside the logical boundaries of 0 and 1 if we use OLS.", correct = T),
  answer("Because the variance of our dependent variable is not large enough."),
  answer("Because we cannot calculate a mean for our dependent variable."),
  answer("Because the standard deviation of our dependent variable equals its mean."),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: Remember what we use OLS for and what our regression line would predict for different kinds of variables."
)
```

```{r q2, echo = FALSE}
question("Why do we need a link function for logistic regression analyses?",
  answer("Because we want to predict probabilities that only range between 0 and 1.", correct = T),
  answer("Because the variance needs to be logged in order to obtain predictions."),
  answer("Because logistic regression models analyze the responses of people participating in surveys.", message = "They do sometimes, but that does not affect the need for a link-function."),
  answer("Because that makes the derivation of the maximum likelihood easier."),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: Consider the special feature of dependent variables in logistic regressions."
)
```

### Coding Exercises

Now let us have a look at the Brexit referendum data again (the data is stored as `df_brexit`). We want to compare some characteristics of the participants in the survey. In a first step, we want you to calculate the average of men in the data set and the average age of women in the data set. Then subtract the average age of women from that of the men. What is the difference in their ages?


```{r exercise1, exercise = TRUE}

```

```{r q3, echo = FALSE}
question_text("What is the age difference between men and women?  Please round to the second digit after the decimal.",
  answer("1.7", correct = T),
  answer("1.70", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: Select the relevant observations with 'filter()' and use the 'mean()' command to calculate the average."
)
```
"
Now calculate the mean age of voters who intend to vote "leave" (`LeaveVote == 1`) and those who intend to vote "remain" (`LeaveVote == 0`).

```{r exercise2, exercise = TRUE}

```

```{r q4, echo = FALSE}
question_text("What is the age difference between leave and remain voters?  Please round to the second digit after the decimal.",
  answer("8.38", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: Select the relevant observations with `filter()` and use the `median()` command to calculate the median."
)
```

Now calculate a logit regression trying to explain `LeaveVote` with `age`, `female`, `edlevel`, `hhincome`, `EuropeanIdentity`, `EnglishIdentity` and `BritishIdentity`. Use the `glm()` command and don't forget to specify the link function. What is the coefficient for `EuropeanIdentity`?  Please round to the second digit after the decimal.

```{r exercise3, exercise = TRUE}

```

```{r q5, echo = FALSE}
question_text("What is the coefficient for `EuropeanIdentity`?  Please round to the second digit after the decimal.",
  answer("-0.80", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: Consider what we did in the lab. The regression command is similar, just add the new independent variables and don't include some that we do not use anymore."
)
```

```{r q6, echo = FALSE}
question("What does the coefficient for `EuropeanIdentity` tell us?",
  answer("That there is a negative relationship with voting Leave", correct = T),
  answer("That there is a positive relationship with voting Leave"),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: Consider the lecture material and what you know about the signs of coefficients."
)
```

### Prediction

Now, let's calculate predicted probabilities. You can use the model below!

```{r , eval = F}
model <- glm(LeaveVote ~ age + female + edlevel + hhincome + EuropeanIdentity +
               EnglishIdentity + BritishIdentity,
             data = df_brexit,
             family = binomial(link = "logit")) 
```

Calculate the predicted probability and the corresponding confidence interval of voting `Leave` for individuals with the highest value for `EuropeanIdentity`, holding all other variables at their means or medians. Remember, that the independent variables (apart from the `age`-variable) are ordinal or dummy variables. Although they are stored as numeric values, the answer options correspond to ordered categorical values. 

What is the predicted probability to vote `Leave` for such voters?Please round to the second digit after the decimal.


```{r exercise4, exercise = TRUE}

```

```{r q7, echo = FALSE}
question_text("  What is the predicted probability to vote `Leave` for such voters? Please round to the second digit after the decimal.",
  answer("0.06", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: Start with the lab code and extend the scenarios to the new variables we included. In case you received a prediction of 0.049/0.05, make sure that you set the questionnaire items to their medians."
)
```

Now, calculate predicted probabilities and the corresponding confidence intervals, this time for the lowest value of `EuropeanIdentity`, holding everything else at its mean or median. What is the predicted probability for those individuals? Again, please round to the second digit after the decimal.

```{r exercise5, exercise = TRUE}

```

```{r q8, echo = FALSE}
question_text("What is the predicted probability for those individuals? Please round to the second digit after the decimal.",
  answer("0.88", correct = T),
  allow_retry = TRUE,
  incorrect = "Hint: Same as before, only the value for `EuropeanIdentity` has to be changed. In case you received a prediction of 0.86, make sure that you set the questionnaire items to their medians."
)
```

```{r q9, echo = FALSE}
question("What do these predictions tell you?",
  answer("People who identify with Europe a lot have a lower probability of voting Leave than those who do not, holding everything else at their respective mean or median.", correct = T),
  answer("People who identify with Europe a lot have a lower probability of voting Leave than those who do not."),
  answer("We cannot say anything about the relationship between high and low identification with Europe and voting Leave because the difference is not statistically significant."),
  answer("People who identify with Europe a lot have a higher probability of voting Leave than those who do not."),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: Compare the predicted probabilities and the uncertainty around them."
)
```


## Submit

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(default_ui(url = ""))
```

