---
title: "Week 04 - Linear Regression Analysis"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: false
    css: "css/learnr-theme.css"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(learnrhash)
gradethis::gradethis_setup(exercise.reveal_solution = FALSE)

library(tidyverse)
library(broom)
library(texreg)
library(qmir22gr)

data(df_opposition) 
```

## Bivariate regression analysis

After you got to know `R` and the `tidyverse` in previous weeks, it's now time to run your first statistical model. Following this week's lab session, you continue wondering why opposition parties in European national parliaments vary in their conflict-seeking behavior. Luckily, we have the data from Tuttnauer (2018) which allows us to explore the behavior of parliamentary opposition parties in 16 EU member states.^[Tuttnauer, O. (2018) "If you can beat them, confront them: Party-level analysis of opposition behavior in European national parliaments". *European Union Politics*, 19(2), 278–298.] 

Before doing so, however, you have to load the required packages. Moreover, you should also familiarize yourself with the data first. 

### Packages

For the following exercises, you'll only need the **tidyverse** package. All other functions should come with base `R`. You should already have `tidyverse` installed, so just go ahead and load it:

```{r load-packages, exercise=TRUE}

```

```{r load-packages-solution}
library(tidyverse)
```

```{r load-packages-check}
grade_this_code("The tidyverse is now loaded!")
```

## Presenting numerical data

Before we start working with the data, let's talk about the presentation of numerical data, first. While we can often estimate values of interest with great accuracy with our statistical models (for example, we might find that the average unemployment rate in our data  is 12.489746168764 %), reporting every digit seems unnecessary (and in this case not very helpful for our overall understanding of the labor market situation). Generally speaking,  we face the challenge of finding a **balance between accuracy and readability** when reporting numerical results. Thus, it is recommendable to report what is often referred to as **"meaningful" or "effective" numbers** --- rounded not too much and not too little.^[Cole, TJ. (2015) "Too many digits: the presentation of numerical data". *Archives of Disease in Childhood* 100. 608-609.]

Yet, what is a "meaningful" number in practice? Unfortunately, there are no universal rules and, to make things even more complicated, the rules for using numbers in academic writing that do exist vary among academic disciplines, journals, and style guides (like the *Chicago Manual of Style* or the *APA Style Manual*). 

For this tutorial (and the weeks to come), you will generally be required to **round numbers to two integer numbers** (integer numbers are positive and negative natural numbers as well as zero: *...-2,-1,0,1,2...*). 

One main exception are numbers with more than two digits *before the decimal sign* (for example, 1564.544). In that case, please provide the whole (rounded) number without digits after the decimal (1565 in this example).

Here are some examples:

| Original value  | Reported value             |
|-----------------|----------------------------|
| 0.23444         | $\rightarrow$ 0.23         |
| 4.021           | $\rightarrow$ 4.0          |
| 4.52861         | $\rightarrow$ 4.5          |
| 13.1248         | $\rightarrow$ 13           |
| 125.1           | $\rightarrow$ 125          |
| 125             | $\rightarrow$ 125          |
| 12,546          | $\rightarrow$ 12546        |


Two more things:

1. As mentioned before, you may sometimes be specifically asked to round to a specified number of digits *after the decimal*. For example, if you are asked to round 125.10243214 to two (integer) digits after the decimal, please provide 125.10 in the solution.

2. **Percentage values** are often stored as values that are bound between 0 and 1 (instead of 0% and 100%). In these cases, you will generally be required to report estimates in **percent**. For example, the vote share stored in a variable in `R` says 0.12 but refers to percentage values. In this case, please report your final result as 12 %.

### Rounding numbers in R

Not surprisingly, we can also round numbers in `R`. The main function to round numeric vectors numbers is called `round()` and takes a numeric vector or value as well as the **number of decimal places** to round to as input: 

- `round(x, digits = 0)`: Rounds the values in its first argument (i.e., x) to the specified number of decimal places.

Here is an example using 189.48615 as input:


```{r round-example}
round(189.48615, 0)
round(189.48615, 1)
round(189.48615, 2)
```

As you can see, you have to be bit careful: If you are asked to report to two meaningful numbers, some people may be inclined to use `round(189.48615, 2)`. However, as  189.48615 has more than two digits *before the decimal sign*, we do not need to report decimal places and `round(189.48615, 0)` is sufficient for our purpose.


Try it yourself! 

**Please round 89.48615 to two meaningful digits**

```{r round-sol,echo = FALSE, eval=FALSE}
round(89.48615, 0)
```


```{r round, exercise = TRUE}

```



```{r round-q, echo = FALSE}
question("Please round 89.48615 to two meaningful digits:",
  answer("89", correct = T, message = "As the number has two integer digits 'before' the decimal, you were right not to report digits 'after' the decimal. Thus, using round(89.48615, 0) is doing the job for you."),
  answer("90", correct = F, message = "You rounded to a single 'significant' number which is not accurate enough. Please report the result rounded to two meaningful digits."),
  answer("89.49", correct = F, message = "Too detailed. You rounded to two digits after the decimal. Please round to two meaningful integer numbers."),
  random_answer_order = TRUE,
  allow_retry = T
)

```




## Getting to know the data

Now it's time to get started with the data. As mentioned before, the data for this tutorial is the same data on opposition parties in national European parliaments that we used in the lab. The data set is already preloaded into the environment and called `df_opposition`. Let's check it out.

### The data set

**How many observations and variables does the data frame have?**

```{r check-data-sol,echo = FALSE, eval=FALSE}
dim(df_opposition)
```


```{r check-data, exercise = TRUE}

```


```{r check-data-q1, echo = FALSE}
question_text("Enter the number of observations here:",
  answer("54", correct = T),
  allow_retry = T,
  incorrect = "Hint: Try dim(), nrow(), or glimpse()."
)

```


```{r check-data-q2, echo = FALSE}
question_text("Enter the number of variables here:",
  answer("17", correct = T),
  allow_retry = T,
  incorrect = "Hint: Try dim(), ncol(), or glimpse()."
)

```

### Dependent variable: Voting against the government
 
As in the lab session, we want to better understand why and how often opposition parties vote against the government. Hence, the  percentage of votes opposed to the government is our dependent variable. The corresponding variable in the data set is called `conflict_p`. 

#### Summary statistics

**What is the median value of this variable (in percent and rounded to two meaningful (integer) digits (e.g., 0.23, 4.5, 13))?**


```{r median-conflict-sol, echo = FALSE, eval=FALSE}
median(df_opposition$conflict_p)
```

```{r median-conflict, exercise = TRUE}

```


```{r median-conflict-q, echo = FALSE}
question("The median  of the conflict measure is (in percent and rounded to two meaningful digits):",
  answer("54", correct = T),
  answer("0.5", correct = F, message = "Please report the result in percent and rounded to two digits."),
  answer("0.54", correct = F, message = "Close but remember that you are supposed to report the median in percent."),
  answer("0.054", correct = F, message = "Please report the result in percent and rounded to two digits."),
  answer("50", correct = F, message = "Please report the result rounded to two meaningful digits."),
  answer("53.68", correct = F, message = "Not wrong per se, but too detailed."),
  random_answer_order = TRUE,
  allow_retry = T
)

```


**What is its standard deviation (in percent and rounded to two meaningful (integer) digits (e.g., 0.23, 4.5, 13))?**

```{r sd-conflict-sol, echo = FALSE, eval=FALSE}
sd(df_opposition$conflict_p) %>% round(2)*100
```

```{r sd-conflict, exercise = TRUE}

```

```{r sd-conflict-q, echo = FALSE}
question_text(
  "The standard deviation of the conflict measure is (in percent and rounded to two meaningful digits):",
  answer("0.27", correct = F, message = "Close, but you were asked to report the result in percent."),
  answer("27", correct = T),
  allow_retry = TRUE
)
```

### Independent variable: Party's seat share

One common argument in the literature on opposition parties is that larger opposition parties oppose the government more often than smaller ones. Let's have a closer look at this hypothesis. As a proxy for party size we are taking its seat share in the parliament stored in the variable `seatshare`. Before we run the regression analysis, however, we should know how the variable looks like.

* **Have a look at the variable using a histogram.**
* **Make the plot look nicer by using `theme_minimal()`.**

```{r histogram, exercise = TRUE}

```

```{r histogram-solution}
ggplot(df_opposition) + 
  geom_histogram(aes(x = seatshare)) + 
  theme_minimal() 
```
 
```{r histogram-code-check}
grade_this_code()
```

<div id="histogram-hint">
**Hint:** Have you (correctly) specified the data and the mappings? The geometric object for creating a histogram is `geom_histogram()`.
</div>

#### Summary statistics

Now that you got a first visual glimpse at the explanatory variable, let's calculate some summary statistics.

**What is the average value of `seatshare` in the data (in percent and rounded to two meaningful digits)?**

```{r avg-seatshare-sol, eval=FALSE, echo = FALSE}
mean(df_opposition$seatshare) %>% round(2)*100
```

```{r avg-seatshare, exercise = TRUE}

```

```{r avg-seatshare-q, echo = FALSE}
question_text(
  "The average value of `seatshare` is (in percent and rounded to two meaningful digits):",
  answer("12", correct = T),
  allow_retry = T
  )
```

**What is the maximum of `seatshare` (in percent and rounded to two meaningful digits)?**

```{r max-seatshare-sol, echo = FALSE, eval=FALSE}
max(df_opposition$seatshare) %>% round(2)*100
```


```{r max-seatshare, exercise = TRUE}

```


```{r max-seatshare-q, echo = FALSE}
question_text(
  "The maximum seat share of an opposition party in the data is (in percent and rounded to two meaningful digits):",
  answer("41", correct = T),
  allow_retry = T
  )
```

## Bivariate regression

Time to run the regression. Regress the average opposition measure (`conflict_p`) on a party's seat share (`seatshare`) using the data set `df_opposition`.

```{r lm_1-sol, eval=FALSE, echo = FALSE}
lm(conflict_p ~ seatshare, data = df_opposition)
```

```{r lm_1, exercise = TRUE}

```

Please provide an interpretation of the estimated regression coefficient. 

*Note that more than one answer can be correct.*

```{r qlm-1-q, echo = FALSE}
question("What does the coefficient estimate tell us about the relationship between seat share and the share of opposing votes to the government?",
  answer("A one unit increase in seat share is associated with an increase of about 56% in votes against the government.", correct = F, message = "Keep in mind that reporting increases in percent is a relative measure. However, the dependent variable is measured in percentage points."),
    answer("A one unit increase in seat share causes an increase in votes against the government of about 56 percentage points.", correct = F, message = "Be careful when making causal statements. Saying increases in seat share cause an increase in votes would only hold if you assume that there are no confounders, i.e. no unobserved factors that affect both the size of a party in parliament and its behavior vis-a-vis the government. Without randomization or further control variables, this seems not realistic to hold in practice."),
  answer("An increase of one percentage point in vote share is associated with a an increase of about 0.56 percentage points in the the share of votes against the government.", correct = T),
    answer("We expect an increase of about 56 percentage points in the share of votes against the government when the seat share variable increases by one unit.", correct = T),
  random_answer_order = TRUE,
  allow_retry = T
)

```

### Predicted values

Not only does the model help us understand how much the share of votes against the government is expected to change when party size increases, but it also allows us to make predictions. 

**According to our model, how often do parties with a seat share of 30% oppose the government on average? Please provide the answer in percent with two significant (non-zero) digits.**

*(Hint: Note that the `predict()` function might be helpful for this task. You can re-run the regression model in the chunk below. Alternatively, the correct regression output is also already stored in an object called `lm_seatshare`)*

```{r lm_predict_sol, eval = F, echo = F}
lm_seatshare <- lm(conflict_p ~ seatshare, data = df_opposition)

predict(lm_seatshare, newdata = tibble(seatshare = .3)) %>% round(2)*100
```

```{r lm_predict-setup}
lm_seatshare <- lm(conflict_p ~ seatshare, data = df_opposition)
```


```{r lm_predict, exercise=TRUE, exercise.setup = "lm_predict-setup"}

```

```{r lm_predict_q, echo = FALSE}
question_text(
  "Parties with a seat share of 30% oppose the government on average in __ percent of votes.",
  answer("66", correct = T),
  answer("65.61", correct = F, message = "That is too detailed. Please round the result to two meaningful digits."),
  answer("0.66", correct = F, message = "Please provide the answer in percent."),
  allow_retry = T
)
```

## Multiple Regression Analysis

Before we start running regressions with multiple explanatory variables, let's recap why we might want to do this in the first place. 

### Warm Up

As discussed in class, one of the main reasons for multiple regression is to address concerns about omitted variable bias. If present, it implies that we systematically miss the true parameter in the population model when we run our statistical model. 

**Choose the correct statement:**

```{r assumptions_ovb, echo = FALSE}
question(sprintf("Is our estimate of $\\beta_1$ biased, if we fail to include a variable that affects our outcome, $y$, but is not correlated with out variable of interest, $x_1$?"),
  answer("Yes.", message = "A causal interpretation of $\\beta_1$ is based on the zero conditional mean assumption. Accordingly, unobserved factors and the variable of interest, $x_1$, are uncorrelated. This assumption is holding in the described scenario."),
  answer("No.", correct = T, message = "Correct. Omitted variable bias only occurs if the omitted variable correlates with both our outcome, $y$, and the variable of interest, $x_1$."),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: Have a look at the slides on omitted variable bias."
)

```


## Getting to know your data

We continue working with the data on opposition parties in national European parliaments from Tuttnauer (2018).^[Tuttnauer, O. (2018) "If you can beat them, confront them: Party-level analysis of opposition behavior in European national parliaments". *European Union Politics*, 19(2), 278–298.] The data set is already preloaded into the environment and called `df_opposition`. You are already quite familiar with the data, so hopefully it won't take long for you to answer the following two questions.

**What is the mean share of opposed votes (`conflict_p`) across all opposition parties and countries (in percent and rounded to two meaningful (integer) digits (e.g., 0.23, 4.5, 13))?**


```{r mean_conflict_sol, echo = FALSE, eval=FALSE}
mean(df_opposition$conflict_p) %>% round(2)*100
```


```{r mean_conflict, exercise = TRUE}

```


```{r mean_conflict_q, echo = FALSE}
question_text("Enter the mean percentage share of opposed votes rounded to two meaningful digits:",
  answer("56", correct = T),
  answer("0.56", correct = F, message = "Provide your answer in percent."),
  answer("0.55", correct = F, message = "Provide your answer in percent and rounded to two digits."),
  answer("56.66", correct = F, message = "Too detailed. It is not really relevant whether the vote share is 0.01 % higher or lower in substantive terms. Therefore it suffices to report two non-zero digits in this case."),
  allow_retry = T
)
```


**Now that we know the average share of votes against the government, how many countries have an average opposition score higher than this mean?**

```{r cntry_conflict_sol, echo = FALSE, eval=FALSE}
avg_opp <- mean(df_opposition$conflict_p)

df_opposition %>% 
  group_by(country) %>% 
  summarise(avg_opposition = mean(conflict_p)) %>% 
  filter(avg_opposition > avg_opp) %>%
  count()
```


```{r cntry_conflict, exercise = TRUE}

```

```{r cntry_conflict_q, echo = FALSE}
question_text("Enter the number of countries with an average opposition score above the mean here:",
  answer("7", correct = T),
  answer("24", message = "You calculated the average opposition scores for parties (and not countries)."),
  allow_retry = T
)
```


## Multiple Regression

In the bivariate regression analysis above, we found that size of a party (`seatshare`) is associated with an increase in how often a party votes against the government (`conflict_p`). However, thinking about this association more closely, you fear that there may be unobserved factors that confound the relationship between the two variables of interest both at the party level as well as at the system level. 

Specifically, you are worried that larger parties are on average ideologically closer to government parties than smaller (fringe) parties. Moreover, it is possible that larger parties have more government experience and are, therefore, likely to behave differently vis-a-vis the government than smaller parties. 

In addition, there may be party-system specific factors varying between countries that affect how party size and opposition to the government are related in our data. For example, the degree to which institutional arrangements allow opposition parties to influence decision-making, the general openness of the competition structure of a party system, the overall importance of the parliament for policy-making, the type of government the opposition parties are facing, as well as how high the threshold is to record voting behavior may be theoretically plausible factors that affect the relationship between party size and opposition behavior.

Luckily, the data set by Tuttnauer (2018) includes variables that measure these potential confounders:

* `abs_lrdiff`: Ideological distance from the government as the absolute difference between a party's left-right position and the size-weighted average position of the government coalition.
* `ge`:  Government-experience value that ranges between 0 if the party has always been in opposition and 1 if the party has always been in government.
* `inst_power`:  Institutional power of the opposition measuring the strength of the committee system on a scale from 0 to 7.
* `y_openness`: Average number of distinct government compositions in the past 10 years.
* `importance`: Index ranging between 0 (low-importance parliament) and 1 (high-importance parliament).
* `govtype_mod`: A dummy (=1) for surplus coalitions, in contrast with minimal-winning ones (=0).
* `record_threshold`:The percentage of MPs needed to call for a recorded vote (ranging between 0 and 0.3).

**Run a regression of the conflict measure on seat share controlling for the above mentioned confounders.**

```{r multi_lm_1_sol, eval=FALSE, echo = FALSE}
model <- lm(conflict_p ~ seatshare + abs_lrdiff + ge + inst_power + 
              y_openness + importance + govtype_mod + record_threshold, data = df_opposition)
knitreg(model)
```

```{r multi_lm_1, exercise = TRUE}

```

**What is the the size of the estimate for seat share (rounded to two meaningful non-integer digits)?**

```{r multi_lm_1_q, echo = FALSE}
question_text(
  "The estimate for seatshare (rounded to two meaningful digits) is:",
  answer("0.57", correct = T),
  answer("57", correct = F, message = "Close, but this time you were actually supposed to report the size of the estimate without transforming it the into percentage share."),
  allow_retry = T
  )
```

### Prediction

Let's use the model to calculate the predicted share of opposed votes for meaningful values that we are interested in. 

**According to our model, how much more/less do parties with a seat share of 10% oppose the government on average in comparison to a party with a seat share of 30% holding all other variables at their mean/median? Please provide the answer in percent with two meaningful (integer) digits.**

*You can re-run the regression model in the chunk below. Alternatively, the correct regression output is also already stored in an object called `out_multivar`.*

```{r multi_predict_sol, eval = F, echo = F}
out_multivar <- lm(conflict_p ~ seatshare + abs_lrdiff + ge + inst_power + 
              y_openness + importance + govtype_mod + record_threshold, data = df_opposition)

df_scenario <- tibble(
   seatshare = c(0.1, 0.3)
  , abs_lrdiff = mean(df_opposition$abs_lrdiff)
  , ge = mean(df_opposition$ge)
  , inst_power = mean(df_opposition$inst_power)
  , y_openness = mean(df_opposition$y_openness)
  , importance = mean(df_opposition$importance)
  , govtype_mod = median(df_opposition$govtype_mod)
  , record_threshold = mean(df_opposition$record_threshold)
)

predict(out_multivar, newdata = df_scenario) %>% round(2)*100
(predict(out_multivar, newdata = df_scenario)[1]-predict(out_multivar, newdata = df_scenario)[2]) %>% round(2)*100
```

```{r multi_predict-setup}
out_multivar <- lm(conflict_p ~ seatshare + abs_lrdiff + ge + inst_power + 
              y_openness + importance + govtype_mod + record_threshold, data = df_opposition)
```

```{r multi_predict, exercise=TRUE, exercise.setup = "multi_predict-setup"}

```

<div id="predict-hint">
**Hint:** The `predict()` function might be helpful for this task. Make sure you only set continuous variables to their mean when specifying the scenarios. Categorical/dummy variables should be set to their median value.
</div>


```{r multi_predict_q, echo = FALSE}
question_text(
  "The difference in the share of votes opposing the government between parties with a seat share of 10% and 30% is __ percentage points.",
  answer("11", correct = T),
  answer("11.46", correct = F, message = "Close, but this is too detailed. Please provide the answer with two meaningful digits only."),
  answer("0.11", correct = F, message = "Please provide the answer in percentage points."),
  answer("12", correct = F, message = "Close. However, it seems as if you rounded the predicted values for the two scenarios first and then calculated the difference. Make sure to round your results only in the final step of your analysis -- so in this case after you calculated the difference."),
  allow_retry = T
)
```

## Interpreting binary variables

Usually, we should be very carefully interpreting control variables by themselves, at least if we didn't spent much time thinking about potential confounders affecting the relationship between a particular control variable and the outcome of interest. For more details on this topic, check out this great overview essay, "The Causal Interpretation of Estimated Associations in Regression Models", by Keele et al. (2019) which is available here: https://doi.org/10.1017/psrm.2019.31


**In this case, however, you are required to make an exception and provide an interpretation of the estimate for `govtype_mod` from the regression model that you ran above.**

Recall how the explanatory variable of interest is coded: It is a dummy variable coded as 1 for surplus coalitions and 0 for minimal-winning coalitions.

*More than one answer can be correct. You can use the following code chunk to re-run the model. Alternatively, the correct regression output is also already stored in an object called `out_multivar`.*

```{r multi_lm2, exercise=TRUE, exercise.setup = "multi_predict-setup"}

```

```{r multi_qlm_1_q, echo = FALSE}
question("What does the estimated regression model tell us about the relationship between government coalition types (`govtype_mod`) and the share of votes opposed to the government?",
  answer("A one unit increase in `govtype_mode` causes a decrease of about 0.04 percentage points in votes against the government, holding everything else contant.", correct = F, message = "Be very careful to make causal statements when using regression analysis without a randomization strategy or thorougly testing for omitted variable bias. Moreover, you are dealing with a dummy variable so it is not wrong per se but generally recommendable to provide interpretations for the respective category values instead of only talking about what happens after one 'unit' increases."),
  answer("Opposition parties facing a surplus coalition government vote on average 0.04 percentage points less often against the government than opposition parties facing a minimal-winning coalition, holding everthing else contant.", correct = T),
  answer("Opposition parties facing a minimal-coalition government vote on average 0.04 percentage points less often against the government than opposition parties facing a surplus coalition government, holding everthing else contant.", correct = F, message = "Make sure that your reference category is correct."),
  answer("Opposition parties facing a minimal-coalition government vote against the government in 83% of the votes on average, setting all other variables to zero.", correct = T, message = "Correct. Good job not neglecting the fact that the intercept has a substantive interpretation, too. It is the expected value when all variables (including `govtype_mod`) are set to zero."),
  random_answer_order = TRUE,
  allow_retry = T
)

```

## Model assumptions

It's important to keep in mind what we are assuming about the models that we run on our data to **infer** what is going on in the real world.

```{r assumptions_causal, echo = FALSE}
question("Which key assumption is required to interpret an estimated relationship between an explanatory variable and the dependent variable as causal?",
  answer("Unobserved factors have an average of zero and do not depend on the explanatory variable.", correct = T),
  answer("The parameters are linear.", message = "This is an important assumption of the  'linear' model that we are building, however, not sufficient for a causal interpretation of the estimates."),
  answer("There is a correlation between the dependent and independent variable.", message = "While correlation is a necessary condition to find a causal effect, it is certainly not sufficient."),
  answer("There is no measurement error in the dependent variable.", message = "A nice feature of regression analysis is that (random) measurement error in the dependent variable does not bias the regression estimates because it does not change the average values of a variable."),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Hint: Have a look at the slides on the zero conditional mean assumption."
)

```


## Submit

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(default_ui(url = "https://ilias.uni-mannheim.de/goto.php?target=svy_1233629&client_id=ILIAS"))
```



